---
title: "DDP Using Leaflet"
author: "Patrick Brooks"
date: "4/12/2020"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(leaflet)
library(maps)
library(dplyr)
library(geojsonio)

# Download the jason version of a US map
#   help iwith US Map https://github.com/rstudio/leaflet/issues/498
states <- geojson_read( 
           x = "https://raw.githubusercontent.com/PublicaMundi/MappingAPI/master/data/geojson/us-states.json", 
           what = "sp")




```

# Project Purpose

The United States Centers for Disease Control (CDC) provides a simple CSV file 
of reported [COVID-19](https://www.cdc.gov/coronavirus/2019-ncov/cases-updates/cases-in-us.html).  cases.  The data table contains the state, recorded cases, the state CDC page.  This exercise will use Leaflet to produce an interactive map of this data.  
The code used in this exercise is based in part on a [leaflet example](https://rstudio.github.io/leaflet/choropleths.html).  

# Methodology
The CDC data contains several US territories not covered by the JSON map.  The CDC data was
manually cleaned for expediency.  The JSON map use contained a dataset with population density.  The variables in the CDC data were added to the dataframe inside JSON map.

* A simple "Blues" pallet was used to color code the states based upon the reported cases.  
* Labels were added and the user can mouse over a state and see cases and the State's CDC site.  
* A legend was added for a macro understanding of the cases reported.  

```{r CDC_Data, echo= FALSE}
d_states = data.frame(readxl::read_xlsx("orgStates Reporting Cases of COVID-19 to CDC.xlsx"))

sp_states = states
# Add CDC data to json map
sp_states$cases = d_states$Cases.Reported[sp_states$name == d_states$Jurisdiction]
sp_states$URL   = d_states$URL[sp_states$name == d_states$Jurisdiction]
bins <- c(0, 500, 1000, 5000, 10000, 50000, 100000, Inf)
pal <- colorBin("Blues", domain = sp_states$cases, bins = bins)

labels <- sprintf(
        "<strong>%s</strong><br/>%g people<br>%s",
        sp_states$name, sp_states$cases,sp_states$URL) %>% lapply(htmltools::HTML)


l_us = leaflet(sp_states) %>% addTiles() %>% setView(-111.656944, 34.29, zoom = 4)


l_us = l_us %>% addPolygons(
        fillColor = pal(sp_states$cases),
        weight = 2,
        opacity = 1,
        color = "white",
        dashArray = "3",
        fillOpacity = 0.7,
        highlight = highlightOptions(
                weight = 5,
                color = "#666",
                dashArray = "",
                fillOpacity = 0.7,
                bringToFront = TRUE),
        label = labels,
        labelOptions = labelOptions(
                style = list("font-weight" = "normal", padding = "3px 8px"),
                textsize = "15px",
                direction = "auto"))

l_us %>% addLegend(pal = pal, values = ~cases, opacity = 0.7, 
                 title = "US Reported Covid19 cases 04/10/2020", position = "bottomright")
         
```

# Summary & code listing

This project demonstrates that creating simple interactive spacial maps can be accomplished
with a a little creative effor.  Below is the code that generated the graphic.


```{r code_inclusion, ref.label=knitr::all_labels(), echo = TRUE, eval = FALSE}

```


